### 2018-01-10 ident/docs/ident_manuscript.tex
complete intro with experimental design focus

Algorithms have been extensively developed for establishing structural identifiability of dynamic models of biological systems \parencite{Audoly2001}(IEEE Trans paper from 2007). Most of these algorithms use methods based on differential algebra \parencite{Ljung1994}. More recent methods take a profile-likelihood-based approach to establish both structural and practical identifiability \parencite{Raue2009a}. All of these methods are heavily dependent on the availability of dynamic time course data for all the output variables of the system and scale poorly with increase in size of the modeled system \parencite{Bellu2007}.

Computational burden due to poor scalability can be partly addressed with the current increases in computational power. Although, improvements in measurement technologies can enable one to obtain high frequency dynamic data, no scalable methodologies exist to utilize the abundantly available steady state data for parameter estimation and identifiability for large metabolic networks.

In this paper, we propose a methodology to establish practical identifiability for parameters in kinetic models of metabolism. We present a computer algebra-based method that can facilitate practical identifiability as well as experimental design for estimating parameters separately for each individual reaction within a metabolic network based on available steady state experimental data. This enables us to address the twin issues of scalability and data availability. We illustrate the utility of this method by applying it for a small network of gluconeogenesis in \textit{E. coli} and demonstrating our ability to propose experiments that will facilitate parameter estimation for a kinetic model of this network. We also demonstrate the scalability of the proposed methodology to facilitate experimental design by applying it to a relatively larger metabolic network of the human red blood cell hepatocyte.

>>>>

Methods and tools for structural identification of parameters based on differential algebra \parencite{Ljung1994, Audoly2001, Bellu2007} and profile likelihood \parencite{Raue2009a} are available. However, only the profile likelihood-based methods enable experimental design by facilitating practical identification of parameters. Nonetheless, these methods still depend on solving a non-convex nonlinear least squares problem (Equation \ref{eq:lsqopt}) to get likelihood estimates of parameters, and hence still suffer from all the inherent difficulties associated with obtaining global optimal solutions for non-convex optimization problems. This also makes them un-scalable for experimental design and practical identifiability of parameters in kinetic models of large metabolic networks.

In this paper, we propose a scalable methodology to establish practical identifiability for parameters in kinetic models of metabolism using abundantly available steady state concentration and flux data. We present a computer algebra-based method that can facilitate practical identifiability as well as experimental design for estimating parameters separately for each individual reaction within a metabolic network based on available steady state experimental data. We illustrate the utility of this method by applying it for a small network of gluconeogenesis in \textit{E. coli} and demonstrating our ability to propose experiments that will facilitate parameter estimation for a kinetic model of this network. We also demonstrate the scalability of the proposed methodology to facilitate experimental design by applying it to a relatively larger metabolic network of the human red blood cell hepatocyte.
